Processing of visual and non-visual naturalistic spatial information in the "parahippocampal place area"

### 2021-01-12 paper/main.tex
restructure event modeling in methods v2

\item at the moment "Annotation, Regressors and Contrasts" follows
    "Preprocessing" (right before "Statistical Analysis" as we agreed on); going
        back and forth a couple of times, imo there is no perfect location (at
        least in case we do not split that part into "Annotation" and
        "Regressors \& Contrasts" again.

\item structure of ``Annotation, Regressors \& Contrasts'' !! -> everything has
    pros and cons (at the moment lots of subheading since easier to navigate);
    at the moment I think (again) it makes more sense to mention AD first since
    it is the focus of the movie and the AV serves as, well, what exactly?

@@

% VIS study picture categories
All stimuli of the VIS study were already used in a previous study
\citep{haxby2011common}.\todo{I guess they only report the ones they
designed themselves in detail? -> check again}
%
There were 24 unique grayscale images (s. Figure 2 in
\citep{sengupta2016extension}) for each of six stimulus categories:
% category details
faces (cropped from original background presented on a medium-gray background),
bodies (female and male without heads in clothes; medium-gray background),
objects (e.g. binoculars, screw, chair, mug; medium-gray background), houses
(medium-gray background) and outdoor scenes comprising of nature and street
scenes, and phase scrambled images.
% controlled luminance and size
Images were matched in luminance and size of 400$\times$\unit[400]{px}
(displayed at a size of approximately
\unit[10]{$^{\circ}$}$\times$\unit[10]{$^{\circ}$} visual angle).
% procedure: presentation & instructions
During the experiment, subjects were presented with four block-design runs, with
two \unit[16]{s} blocks per stimulus category in each run.
% temporal progression
Each image was shown for \unit[900]{ms} and images were separated by a fixation
cross lasting \unit[100]{ms}.
% task
Subjects also performed a one-back matching task to keep them attentive.

@@

% traditionally
Traditionally, researchers specifically design stimuli that presumably would
trigger the perceptual process of interest while also controlling confounding
variables (e.g. color, luminance).
% our approach: annotation and regressors
Naturalistic stimuli [might] more closely mimic how we experience the world
around us but have the drawback that they are usually fix and that their
temporal structure of embedded stimulus features is unknown.\todo{phrasing}
% annotation ftw
Hence, we annotated a variety of stimulus features that later served as a basis
(for GLM regressors) to model hemodynamic responses.

\subsubsection{Events (based on annotations)}

\todo[inline]{mention low-level confound annotations here or (as it is now) in
modeling}

@@

% irrespective of visual content
Notably, categorization was performed irrespectively of a movie frame's precise
content and thus without controlling confounding variables (e.g. presence of
faces).
% AV COPEs: intro
The rationale that this approach would still yield categories correlating with
increased hemodynamic activity in the PPA was twofold.
% cinematographic rationale cut to new scenes: establishing shots
The cinematographic rationale was that movies tend to establish the setting
(\texttt{vse\_new}) and the spatial relationships within a setting
(\texttt{vpe\_new}) at the beginning of a movie.
% cut to recurrent scene
Later in the movie, the field sizes of movie shots tend to decrease when
switching back to already established settings (\texttt{vse\_old}).
% similarly for shots within a scene
Moreover, shots within a recurring locale or setting tend to shift to depicting
persons and objects (\texttt{vpe\_old}) more relevant to the evolved plot
\citep{brown2012cinematography, mercado2011filmmakers}.
% neuroscientific rationale
The neuroscientific rationale was that activity in the PPA is greater when
participants view novel versus repeated scenes or view-points
\citep{epstein1999parahippocampal, grill2006repetition}.

@@

\paragraph{Audio-description}

@@

% AD annotation
For the analysis of the AD stimulus, we extended a publicly available annotation
of speech \citep{haeusler2020speechanno} by further annotating words spoken by
the audio-description's narrator.
% annotation procedure
Two persons performed a categorization of nouns that the narrator uses to
describe the movie's missing visual content based on a set of rules.
% reference to table with rules and examples
A complete overview of all 18 noun categories, the rule for a word to belong to
a category, and examples can be seen in Table \ref{tab:descr-nouns-rules}.
% preliminary annotation
A preliminary annotation was performed by one person according to the rules.
% corrections
Minor corrections due to incorrectly applied rules were done by the author.
% why and how of categories 1
Some categories reflect the verbal counterpart of the stimulus categories that
were used in the visual localizer experiment (e.g. \texttt{body}, \texttt{face},
\texttt{head}, \texttt{object}, \texttt{setting\_new}, and
\texttt{setting\_rec}).
% why and how of categories 2
Other categories were created to semantically cluster remaining nouns into
categories that had no counterpart in the visual localizer experiment (e.g.
\texttt{bodypart}, \texttt{female}, \texttt{fname}, \texttt{furniture},
\texttt{geo}, \texttt{geo-room}, \texttt{male}, and \texttt{persons}).
% categories in detail
%Nouns were categorized by the verbal clue they provide about the
%cinematographic scene's environment (\texttt{geo}, \texttt{geo-room};
%\texttt{setting\_new}, \texttt{setting\_old}), its inherent persons (e.g.
%\texttt{female}, \texttt{male}, \texttt{persons}), a person's body or worn
%clothes (\texttt{face}, \texttt{head}, \texttt{body}, \texttt{bodypart}), and a
%scene's inherent objects (\texttt{object}, \texttt{furniture}).  problem with
%hierarchical categories
The categories \texttt{setting\_new} and \texttt{setting\_rec} comprise not just
words that describe a setting as a whole (e.g. ``[in] Greenbow'', ``[in a]
disco'', ``[the platoon wades through a] rice field''). They also comprise words
that could count as a member of another categories in case the narrator uses
these words to indicate a switch from one setting to another (e.g. ``[a]
physician'').
% how it was handled
These nouns were flagged with both category during the initial annotation
procedure. In context of the current analyses these word are considered to
belong exclusively to the higher-level category of changing a setting
(\texttt{se\_new} or \texttt{se\_old}).\todo{if you know what I mean}
% pooling of categories
Lastly, some categories that were semantically similar categories that offer
only a small amount of counts were pooled resulting in 11 final categories (see
Table \ref{tab:ao-events}) that were used as event categories [to model
different hemodynamic responses].

@@

\todo[inline]{SE: Table 2 ist relativ kryptisch bzw für jemanden, der sich zum
    ersten Mal damit auseinander setzt kaum zu verstehen. Auch ist der Bezug zum
    GLM unklar, d.h., was davon wird verwendet und was verbirgt sich dahinter;
    COH: I dropped the table of counts per segment. The information is
    redundant since the counts are not the counts of events of the actual
regressors (cf. Table \ref{tab:av-events})}

\begin{comment}
% nouns counts
The resulting counts for the whole stimulus and the eight stimulus segments used
during fMRI scanning can be seen in Table \ref{tab:descr-nouns-counts}.

\begin{table*}[h!]
    \caption{Descriptive nouns spoken by the audio-descriptions narrator:
        counts for the whole audio-only stimulus and its eight segments used
        during fMRI scanning.}
\label{tab:descr-nouns-counts}
\begin{tabular}{llllllllll}
\toprule
\textbf{category} & \textbf{all} & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8} \\
\midrule
body & \anBodyAll & \anBodyI & \anBodyII & \anBodyIII & \anBodyIV & \anBodyV & \anBodyVI & \anBodyVII & \anBodyVIII \tabularnewline
bodypart &  \anBodypartAll & \anBodypartI & \anBodypartII & \anBodypartIII & \anBodypartIV & \anBodypartV & \anBodypartVI & \anBodypartVII & \anBodypartVIII \tabularnewline
face & \anFaceAll & \anFaceI & \anFaceII & \anFaceIII & \anFaceIV & \anFaceV & \anFaceVI & \anFaceVII & \anFaceVIII \tabularnewline
female & \anFemaleAll & \anFemaleI & \anFemaleII & \anFemaleIII & \anFemaleIV & \anFemaleV & \anFemaleVI & \anFemaleVII & \anFemaleVIII \tabularnewline
females & \anFemalesAll & \anFemalesI & \anFemalesII & \anFemalesIII & \anFemalesIV & \anFemalesV & \anFemalesVI & \anFemalesVII & \anFemalesVIII \tabularnewline
fname & \anFnameAll & \anFnameI & \anFnameII & \anFnameIII & \anFnameIV & \anFnameV & \anFnameVI & \anFnameVII & \anFnameVIII \tabularnewline
furniture & \anFurnitureAll & \anFurnitureI & \anFurnitureII & \anFurnitureIII & \anFurnitureIV & \anFurnitureV & \anFurnitureVI & \anFurnitureVII & \anFurnitureVIII \tabularnewline
geo & \anGeoAll & \anGeoI & \anGeoII & \anGeoIII & \anGeoIV & \anGeoV & \anGeoVI & \anGeoVII & \anGeoVIII \tabularnewline
geo-room & \anGeoroomAll & \anGeoroomI & \anGeoroomII & \anGeoroomIII & \anGeoroomIV & \anGeoroomV & \anGeoroomVI & \anGeoroomVII & \anGeoroomVIII \tabularnewline
head & \anHeadAll & \anHeadI & \anHeadII & \anHeadIII & \anHeadIV & \anHeadV & \anHeadVI & \anHeadVII & \anHeadVIII \tabularnewline
male & \anMaleAll & \anMaleI & \anMaleII & \anMaleIII & \anMaleIV & \anMaleV & \anMaleVI & \anMaleVII & \anMaleVIII \tabularnewline
males & \anMalesAll & \anMalesI & \anMalesII & \anMalesIII & \anMalesIV & \anMalesV & \anMalesVI & \anMalesVII & \anMalesVIII \tabularnewline
mname & \anMnameAll & \anMnameI & \anMnameII & \anMnameIII & \anMnameIV & \anMnameV & \anMnameVI & \anMnameVII & \anMnameVIII \tabularnewline
object & \anObjectAll & \anObjectI & \anObjectII & \anObjectIII & \anObjectIV & \anObjectV & \anObjectVI & \anObjectVII & \anObjectVIII \tabularnewline
objects & \anObjectsAll & \anObjectsI & \anObjectsII & \anObjectsIII & \anObjectsIV & \anObjectsV & \anObjectsVI & \anObjectsVII & \anObjectsVIII \tabularnewline
persons & \anPersonsAll & \anPersonsI & \anPersonsII & \anPersonsIII & \anPersonsIV & \anPersonsV & \anPersonsVI & \anPersonsVII & \anPersonsVIII \tabularnewline
setting\_new & \anSettingnewAll & \anSettingnewI & \anSettingnewII & \anSettingnewIII & \anSettingnewIV & \anSettingnewV & \anSettingnewVI & \anSettingnewVII & \anSettingnewVIII \tabularnewline
setting\_rec & \anSettingrecAll &
\anSettingrecI & \anSettingrecII & \anSettingrecIII & \anSettingrecIV & \anSettingrecV & \anSettingrecVI & \anSettingrecVII & \anSettingrecVIII \tabularnewline
% ++ & \anAll & \anI & \anII & \anIII & \anIV & \anV & \anVI & \anVII & \anVIII \tabularnewline
\bottomrule
\end{tabular}
\end{table*}

\end{comment}

@@

\subsubsection{Modeling}

\paragraph{audio-description: regressors}

% modeling as boxcar function
Events were were modeled as boxcar function from onset to offset of each word.

\todo[inline]{Gab es ein separates null-event für jeden zeitraum zwischen zwei
annotierten events oder war das ein regressor der immer wieder kam? COH: uhm,
sorry, what? the following explanation must be misleading?}
% null regressors
If a stimulus segment did not contain an event of a category, a null regressor
was chosen for that category in that segment.\todo{wenn ich "null regressor"
    nehme, darf ich den ganzen Run, glaube ich, in der 2nd level Analyse gar
    nicht benutzen?  Hatte, glaube ich, Jeanette Mumford in einem ihren Videos
mal gesagt}

% categories taken from movie cut annotation
To create contrasts of cross-modal negative control and to capture variance of a
possibly confounding change of the soundscape after a cut, we also implemented
the five movie cut-related regressors used in the GLM of the AV stimulus (see
details below) in the model of the AD stimulus.
% nuisance regressors
\todo[inline]{"Continuous bins lasting 40ms?"; COH: goddamnit! I've rewritten
    this explanation a lot and finally stuck to the phrasing written in
    \citep{wilson2008beyond} who did the same; it is rephrased now though}
%
Lastly, we created two nuisance regressors based on left-right difference in
volume and root mean square energy averaged across the length of every movie
frame (\unit[40]{ms}) to capture variance correlating with low-level perceptual
processes.
% Reference to table
An overview and the respective counts of the eleven speech-related event
categories, the five movie cut-related categories, and the two low-level
confounds can be found in Table \ref{tab:ao-events}.
% convolving
Events for all regressors of the AV stimulus were convolved with a
double gamma hemodynamic response function (HRF).

\paragraph{movie: regressors}

\todo[inline]{besser erklären (was heißt "unfamilar perspective"? Ab wann ist
ein setting "familiar"?; COH: should be clearer now; maybe rephrase into a
couple of sentences to improve readability}

% null regressor
As for the regressors for AD stimulus, a null regressor was chosen for a
regressor in a segment if that segment did not contain an event of the
corresponding category.
% no cut condition: intro
A sixth regressors of pseudo-randomly chosen frames (\texttt{no\_cut}) was
created as a control condition to create negative control contrasts:
% procedure
Frames were drawn from movie shots lasting at least \unit[20]{s} and had to have
a minimum distance of \unit[10]{s} to a movie cut and to another
\texttt{no\_cut} event.

\todo[inline]{mention low-level feature events for AV here as well and not just
in the figure caption}

% reference to table
A short explanation and counts of these six categories modeled as impulse
events, and six additional event categories for low-level auditory and visual
features of the movie can be found in Table \ref{tab:av-events}.
% convolving
Events for all regressors of the AV stimulus were convolved with a
double gamma hemodynamic response function (HRF).

% temporal derivatives
Temporal derivatives were included in the design matrix to compensate for
regional differences between modeled and actual HRF \citep{friston1998event}.
% motion parameters
Six motion parameters were also used as additional nuisance regressors.
% high-pass filtering
Finally, designs were filtered with the same high-pass filter (cutoff period of
\unit[150]{s}) as the BOLD time series.

@@

@@

>>>>

\item confound regressors von modeling hoch zu eventes

\item revise captions of Table 1 + 3

@@

% VIS study picture categories
All stimuli of the VIS study were already used in a previous study
\citep{haxby2011common}.\todo{I guess they only report the ones they
designed themselves in detail? -> check again}
%
There were 24 unique grayscale images (s. Figure 2 in
\citep{sengupta2016extension}) for each of six stimulus categories:
% category details
faces (cropped from original background presented on a medium-gray background),
bodies (female and male without heads in clothes; medium-gray background),
objects (e.g. binoculars, screw, chair, mug; medium-gray background), houses
(medium-gray background) and outdoor scenes comprising of nature and street
scenes, and phase scrambled images.
% controlled luminance and size
Images were matched in luminance and size of 400$\times$\unit[400]{px}
% procedure: presentation & instructions
During the experiment, subjects were presented with four block-design runs, with
two \unit[16]{s} blocks per stimulus category in each run.
% temporal progression
Each image was shown for \unit[900]{ms} and images were separated by a fixation
cross lasting \unit[100]{ms}.
% task
Subjects also performed a one-back matching task to keep them attentive.

@@

% traditionally
Traditionally, researchers specifically design stimuli that presumably would
trigger the perceptual process of interest while also controlling confounding
variables (e.g. color, luminance).
% our approach: annotation and regressors
Naturalistic stimuli [might] more closely mimic how we experience the world
around us but have the drawback that they are usually fix and that their
temporal structure of embedded stimulus features is unknown.
% annotation ftw
Hence, we annotated a variety of stimulus features that later served as a basis
(for GLM regressors) to model hemodynamic responses.

\subsubsection{Events}

@@

% irrespective of visual content
Notably, categorization was performed irrespectively of a movie frame's precise
content and thus without controlling confounding variables (e.g. presence of
faces).
% AV COPEs: intro
The rationale that this approach would still yield categories correlating with
increased hemodynamic activity in the PPA was twofold.
% cinematographic rationale cut to new scenes: establishing shots
The cinematographic rationale was that movies tend to establish the setting
(\texttt{vse\_new}) and the spatial relationships within a setting
(\texttt{vpe\_new}) at the beginning of a movie.
% cut to recurrent scene
Later in the movie, the field sizes of movie shots tend to decrease when
switching back to already established settings (\texttt{vse\_old}).
% similarly for shots within a scene
Moreover, shots within a recurring locale or setting tend to shift to depicting
persons and objects (\texttt{vpe\_old}) more relevant to the evolved plot
\citep{brown2012cinematography, mercado2011filmmakers}.
% neuroscientific rationale
The neuroscientific rationale was that activity in the PPA is greater when
participants view novel versus repeated scenes or view-points
\citep{epstein1999parahippocampal, grill2006repetition}.
% no cut condition: intro
In order to create a control condition, a sixth category (\texttt{no\_cut}) was
created by pseudo-randomly drawing movie frames from continuous movie shots that
lasted longer than \unit[20]{s}.
% no cuts: how
Events were drawn from movie shots in such a way that each event had a minimum
temporal distance of at least \unit[10]{s} to a movie cut and to another
\texttt{no\_cut} event.

\paragraph{AV confounds}

% visual confounds -> cf. table

% audio-track confounds
Lastly, we created two nuisance regressors based on left-right difference in
volume and root mean square energy averaged across the length of every movie
frame (\unit[40]{ms}) to capture variance correlating with low-level perceptual
processes.

\todo[inline]{SE: Continuous bins lasting 40ms?}

@@

\paragraph{audio-description}

@@

% AD annotation
For the analysis of the AD stimulus, we extended a publicly available annotation
of speech \citep{haeusler2020speechanno} by further annotating words spoken by
the audio-description's narrator.
% annotation procedure
Two persons performed a categorization of nouns that the narrator uses to
describe the movie's missing visual content based on a set of rules.
% preliminary annotation
A preliminary annotation was performed by one person according to the rules.
% corrections
Minor corrections due to incorrectly applied rules were done by the author.
% reference to table with rules and examples
A complete overview of all 18 noun categories, the rule for a word to belong to
a category, and examples can be seen in Table \ref{tab:descr-nouns-rules}.
% why and how of categories 1
Some categories reflect the verbal counterpart of the stimulus categories that
were used in the visual localizer experiment (e.g. \texttt{body}, \texttt{face},
\texttt{head}, \texttt{object}, \texttt{setting\_new}, and
\texttt{setting\_rec}).
% why and how of categories 2
Other categories were created to semantically cluster remaining nouns into
categories that had no counterpart in the visual localizer experiment (e.g.
\texttt{bodypart}, \texttt{female}, \texttt{fname}, \texttt{furniture},
\texttt{geo}, \texttt{geo-room}, \texttt{male}, and \texttt{persons}).
% categories in detail
%Nouns were categorized by the verbal clue they provide about the
%cinematographic scene's environment (\texttt{geo}, \texttt{geo-room};
%\texttt{setting\_new}, \texttt{setting\_old}), its inherent persons (e.g.
%\texttt{female}, \texttt{male}, \texttt{persons}), a person's body or worn
%clothes (\texttt{face}, \texttt{head}, \texttt{body}, \texttt{bodypart}), and a
%scene's inherent objects (\texttt{object}, \texttt{furniture}).  problem with
%hierarchical categories
The categories \texttt{setting\_new} and \texttt{setting\_rec} comprise not just
words that describe a setting as a whole (e.g. ``[in] Greenbow'', ``[in a]
disco'', ``[the platoon wades through a] rice field''). They also comprise words
that could count as a member of another categories in case the narrator uses
these words to indicate a switch from one setting to another (e.g. ``[a]
physician'').
% how it was handled
These nouns were flagged with both categories during the initial annotation
procedure. In context of the current analyses these word are considered to
belong exclusively to the higher-level category of changing a setting
(\texttt{se\_new} or \texttt{se\_old}).
% pooling of categories
Lastly, some noun categories that were semantically similar and offered only a
small amount of counts were pooled resulting in 11 final event categories (see
Table \ref{tab:ao-events}). These event categories were then used to model
hemodynamic responses.

\paragraph{AD confounds}

% audio confounds -> cf. AV & table

% final reference to table
An overview and the respective counts of the eleven speech-related event
categories, the five movie cut-related categories, and the two low-level
confounds can be found in Table \ref{tab:ao-events}.

@@

@@

\subsubsection{Hemodynamic Modeling}

% movie events
The events of the movie cut related categories and \texttt{no\_cut} events were
modeled as impulse events [lenght = 20 ms?].\todo[actually, that might not be
true}

AV:
% convolving

AO:
% modeling as boxcar function
Events were were modeled as boxcar function from onset to offset of each word.

\todo[inline]{SE: gab es ein separates null-event für jeden zeitraum zwischen
zwei annotierten events oder war das ein regressor der immer wieder kam?}

% AD: null regressors
If a stimulus segment did not contain an event of a category, a null regressor
was chosen for that category in that segment.
% AV: null regressor
As for the regressors for AD stimulus, a null regressor was chosen for a
regressor in a segment if that segment did not contain an event of the
corresponding category.

% convolving
Events for all regressors of the AV and AO timulus were convolved with a
double gamma hemodynamic response function (HRF).
% temporal derivatives
Temporal derivatives were included in the design matrix to compensate for
regional differences between modeled and actual HRF \citep{friston1998event}.
% motion parameters
Six motion parameters were also used as additional nuisance regressors.
% high-pass filtering
Finally, designs were filtered with the same high-pass filter (cutoff period of
\unit[150]{s}) as the BOLD time series.

@@

\todo[inline]{regressors of other stimulus in GLM of other stimulus}
% categories taken from movie cut annotation
Maybe only say in table caption: To create contrasts of cross-modal negative
control and to capture variance of a possibly confounding change of the
soundscape after a cut, we also implemented the five movie cut-related
regressors used in the GLM of the AV stimulus (see details below) in the model
of the AD stimulus.

@@

\paragraph{included regressors}

% AO events in AV

% AV events in AO
